---
title: "TeamProject2"
author: "Dance Dance Revolution: Mario Prata Dias Alvarez, Sophie Kate Edelman, Meghan Sarah Hoffman, Elias Florentin John, Kritika Singh, Benjamin Victor Tward"
date: "2025-10-23"
output: 
  html_document:
    toc: True
    toc_float: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
library(caret)
#install.packages("recipes")
library(recipes)
```

## Step 0: Why?

Step 0: Context, Objective, and Financial Model

Project Context: Realistic Spotify Metrics

To ground this analysis in real-world performance, we use approximate public and industry figures for a large streaming platform (Q3 2024 estimates):

```{r}
context_data <- data.frame(
  Metric = c("Monthly Active Users (MAU)", "Premium Subscribers", "Average Revenue Per User (ARPU)", "Estimated Monthly Churn Rate", "Customer Lifetime Value (CLV)"),
  Value = c("~600 Million", "~250 Million (42% of MAU)", "~$4.50 / month", "~2.0%", "~$225"),
  Rationale = c("Based on publicly reported figures.", "Based on publicly reported subscription rates.", "Blended rate across different tiers and regions.", "Industry estimate for major streaming services.", "Calculated as ARPU / Churn Rate = $4.50 / 0.02.")
)
colnames(context_data) <- c("Metric", "Estimated Value", "Rationale")
knitr::kable(context_data, caption = "Realistic Spotify Business Metrics", format = "markdown")
```

Objective

Objective: To maximize user retention and daily active listening time, how can Spotify's recommendation engine use a song's audio features (danceability, energy, valence), track_genre, and other variables to build a predictive model that more accurately identifies high-engagement tracks (i.e., high future popularity) for its personalized playlists like 'Discover Weekly'?

Assumption & Financial Model

Assumption: A $10\%$ increase in the number of high-engagement songs a user hears in their 'Discover Weekly' playlist will lead to a very marginal $0.01\%$ decrease in their probability of churning (canceling their subscription) that month.

We translate this into the following Net Financial Impact model to guide the Level 2 Stacker. The model is designed to maximize Net Impact (Revenue - Loss). The $\$$ values reflect the estimated scaled impact on CLV and promotion costs per decision.

```{r}
financial_data <- data.frame(
  Outcome = c("True Positive (TP)", "False Positive (FP)", "False Negative (FN)"),
  Metric = c("Correctly identifies a Hidden Gem (Target = 'Yes').", "Predicts 'Yes', but actual is 'No'. (Wasteful promotion)", "Predicts 'No', but actual is 'Yes'. (Lost opportunity)"),
  Financial_Value = c("+$200", "-$500", "-$100"),
  Financial_Rationale = c("Revenue: Estimated net monthly value derived from the 0.01% reduction in churn for a high-value, retained customer. (Maximize)", "Cost: Significant loss from license fees, content storage, and opportunity cost of promoting non-engaging content that accelerates churn. (Minimize - Critical Error)", "Cost: Minor loss from missing a viable track and losing the opportunity for increased engagement and retention. (Minimize - Minor Error)")
)
colnames(financial_data) <- c("Outcome", "Metric", "Financial Value", "Financial Rationale")
knitr::kable(financial_data, caption = "Net Financial Impact Model (Per Prediction)", format = "markdown")
```

The objective is to maximize the Net Impact, calculated as:

\text{Net Impact} = (\text{TP} \times \$200) - (\text{FP} \times \$500) - (\text{FN} \times \$100) \## Stacked Model Architecture

The process uses a **6-model Level 1 (L1) Ensemble** feeding into a **Level 2 (L2) Stacker**.

| Level | Model Name | Role | Optimization |
|:-----------------|:-----------------|:-----------------|:-----------------|
| **Level 1** | LR, KNN, ANN, RF, DT, SVM | **Specialists:** Trained on raw features (danceability, genre, etc.) to capture diverse linear and non-linear patterns. | Standard statistical metrics (Kappa). |
| **Level 2** | Cost-Sensitive Logistic Regression | **Producer:** Trained on L1 probability outputs using **weighted learning** based on the $5:1$ loss ratio ($-\$500$ FP cost vs. $-\$100$ FN cost). | Maximize Net Financial Impact by aggressively minimizing False Positives. |

## Level 1 Step 1, 2, 3: Importing, Cleaning, Splitting, and Normalizing the Data
```{r}
source("modules/prep.R")
```


## Level 1 Step 4, 5, 6: Building, Predicting, and Analyzing the Models

```{r}
source("modules/prep.R")
train = data.frame(X_train, popularity = y_train)
test = data.frame(X_test, popularity = y_test)
```

### GLM

Work for this model was done in "modules/ann.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
glm_prob <- readRDS("probabilities/GLMProbs.RDS")
glm_pred <- readRDS("predictions/GLMPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
glm_confusion <- confusionMatrix(as.factor(glm_pred), as.factor(test$popularity), positive = "1")
glm_confusion
```

**What we changed to get the best results**

For the best results, we manipulated the threshold for the predictions to be popular or unpopular. Initially, we used a threshold of 0.5, but after optimizing for accuracy and sensitivity, we used a threshold of -0.6. 

### ANN

Work for this model was done in "modules/ann.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
ann_prob <- readRDS("probabilities/ANNProbs.RDS")
ann_pred <- readRDS("predictions/ANNPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
ann_confusion <- confusionMatrix(as.factor(ann_pred), as.factor(test$popularity), positive = "1")
ann_confusion
```

**What we changed to get the best results**
For the best results, we manipulated the threshold for the predictions to be popular or unpopular. Initially, we used a threshold of 0.5, but after optimizing for accuracy and sensitivity, we used a threshold of 0.4. 

### SVM

Work for this model was done in "modules/svm.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
svm_prob <- readRDS("probabilities/SVMProbs.RDS")[, 1]
svm_pred <- readRDS("predictions/SVMPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
svm_confusion <- confusionMatrix(as.factor(svm_pred), as.factor(test$popularity), positive = "1")
svm_confusion
```

**What we changed to get the best results**
For the best results, we manipulated the threshold for the predictions to be popular or unpopular. Initially, we used a threshold of 0.5, but after optimizing for accuracy and sensitivity, we used a threshold of 0.3. 


### Random Forest

Work for this model was done in "modules/forest.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
forest_prob <- readRDS("probabilities/ForestProbs.RDS")
forest_pred <- readRDS("predictions/ForestPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
forest_confusion <- confusionMatrix(as.factor(forest_pred), as.factor(test$popularity), positive = "1")
forest_confusion
```

**What we changed to get the best results**
For the best results, we manipulated the threshold for the predictions to be popular or unpopular. Initially, we used a threshold of 0.3, but after optimizing for accuracy and sensitivity, we used a threshold of 0.35. 

### KNN

Work for this model was done in "modules/knn.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
knn_prob <- readRDS("probabilities/KNNProbs.RDS")
knn_pred <- readRDS("predictions/KNNPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
knn_confusion <- confusionMatrix(as.factor(knn_pred), as.factor(test$popularity), positive = "1")
knn_confusion
```

**What we changed to get the best results**

### Decision Tree

Work for this model was done in "modules/knn.R", however, since the file sizes were so large, the models, probabilities, and predictions were exported into separate files to be used below.

```{r}
decision_prob <- readRDS("probabilities/DecisionProbs.RDS")
decision_pred <- readRDS("predictions/DecisionPredictions.RDS")
```

To understand the effect of model, we put it through a confusion matrix

```{r}
library(caret)
decision_confusion <- confusionMatrix(as.factor(decision_pred), as.factor(test$popularity), positive = "1")
decision_confusion
decision_confusion$overall["Accuracy"]
```

**What we changed to get the best results**

### All of the Level 1 Models

| Model | Accuracy | Sensitivity | Costs |
|----------------------|----------------|-------------------|----------------|
| GLM | `r glm_confusion$overall["Accuracy"]` | `r glm_confusion$byClass["Sensitivity"]` |  |
| ANN | `r ann_confusion$overall["Accuracy"]` | `r ann_confusion$byClass["Sensitivity"]` |  |
| SVM | `r svm_confusion$overall["Accuracy"]` | `r svm_confusion$byClass["Sensitivity"]` |  |
| Random Forest | `r forest_confusion$overall["Accuracy"]` | `r forest_confusion$byClass["Sensitivity"]` |  |
| KNN | `r knn_confusion$overall["Accuracy"]` | `r knn_confusion$byClass["Sensitivity"]` |  |
| Decision Tree | `r decision_confusion$overall["Accuracy"]` | `r decision_confusion$byClass["Sensitivity"]` |  |

### Combining the models

```{r}
combined_models <- data.frame(ann_pred, decision_pred, forest_pred, glm_pred, knn_pred, svm_pred, test$popularity)
```

## Part 2: Building a Second Level Decision Tree

Since we combined the models right above, there is no need to import of clean them. 

### Step 3: Split the Data

```{r}
trainfrac2 <- 0.7
set.seed(12345)

trainrows <- sample(1:nrow(combined_models), trainfrac2*nrow(combined_models))

train2 <- combined_models[trainrows, ]
test2 <- combined_models[-trainrows, ]
```

### Step 4: Building the Model

```{r}
library(C50)
decision_2 <- C5.0(as.factor(test.popularity) ~., data = train2)
```

### Step 5: Predicting with the Model
```{r}
decision_p2 <- predict(decision_2, test2, type = "prob")
decision_p2 <- decision_p2[, 2]

decision_pred2 <- ifelse(decision_p2 >= .5, 1, 0)
```

### Step 6: Analyzing the Model

```{r}
library(caret)
decision2_confusion <- confusionMatrix(as.factor(decision_pred2), as.factor(test2$test.popularity), positive = "1")
decision2_confusion
```

## Step 7: Conclusion

| Model | Accuracy | Sensitivity | Costs |
|----------------------|----------------|-------------------|----------------|
| GLM | `r glm_confusion$overall["Accuracy"]` | `r glm_confusion$byClass["Sensitivity"]` |  |
| ANN | `r ann_confusion$overall["Accuracy"]` | `r ann_confusion$byClass["Sensitivity"]` |  |
| SVM | `r svm_confusion$overall["Accuracy"]` | `r svm_confusion$byClass["Sensitivity"]` |  |
| Random Forest | `r forest_confusion$overall["Accuracy"]` | `r forest_confusion$byClass["Sensitivity"]` |  |
| KNN | `r knn_confusion$overall["Accuracy"]` | `r knn_confusion$byClass["Sensitivity"]` |  |
| Decision Tree | `r decision_confusion$overall["Accuracy"]` | `r decision_confusion$byClass["Sensitivity"]` |  |
| **2nd Level Decision Tree** | `r decision2_confusion$overall["Accuracy"]` | `r decision2_confusion$byClass["Sensitivity"]` |  |


### Overall Conclusion

### Possible Errors
